TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-06 paper/draft.tex
minor details

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations (e.g. \kepler, \tess), and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches, making detailed and precise knowledge of M dwarf chemical compositions key to constraining the atmospheric conditions of potentially habitable exoplanets.

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade have dramatically increased the sample of known M dwarfs, and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level (e.g. SDSS \citealt{West:2011}; LAMOST \citealt{Yi:2014} and \citealt{Guo:2015}).
The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy, yet pipeline measurements \citep{Perez:2016} of $\teff$ and $\feh$ have only been determined precise to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K \citep{Schmidt:2016}.

Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods: photometric calibrations (\citealt{Johnson:2009}; \citealt{Schlaufman:2010}) or spectroscopic calibrations (\citealt{Terrien:2012}; \citealt{Rojas-Ayala:2012}; \citealt{Mann:2013a}; \citealt{Newton:2014}) with FGK pairs. Our approach is the latter.

>>>>

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations (e.g. \kepler, \tess), and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches, making detailed and precise knowledge of M dwarf chemical compositions key to constraining the formation scenarios and atmospheric conditions of potentially habitable exoplanets.

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade have dramatically increased the sample of known M dwarfs, and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level (e.g. SDSS \citealt{West:2011}; LAMOST \citealt{Yi:2014} and \citealt{Guo:2015}).
The launch of the (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy, yet pipeline measurements \citep{Perez:2016} of $\teff$ and $\feh$ have only been determined precise to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K using atmospheric models \citep{Schmidt:2016}.

\color{gcolor}Why is there a problem determining M dwarf metallicities (and especially for high res \apogee\ spectra)? Emphasize issues: (1) models dont accurately reproduce data; (2) empirical standards are relative not absolute, and they are sparesely sampled. 
Methods for determining metallicity of M dwarfs: (1) from atmospheric models of M dwarfs, (2) empirical methods: photometric calibrations (\citealt{Johnson:2009}; \citealt{Schlaufman:2010}) or spectroscopic calibrations (\citealt{Terrien:2012}; \citealt{Rojas-Ayala:2012}; \citealt{Mann:2013a}; \citealt{Newton:2014}) with FGK pairs. Our approach is the latter.\color{black}
